---
title: "Exploring parks data"
author: "Nicholas Tierney"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libs, echo = FALSE, include = FALSE}
library(tidyverse)
library(vctrs)
library(tsibble)
conflicted::conflict_prefer("filter", "dplyr")
conflicted::conflict_prefer("select", "dplyr")
conflicted::conflict_prefer("pluck", "purrr")
```


# Reading

```{r}
park_visits <- readr::read_csv(here::here("data/park-visits.csv"))
gas_price <- readr::read_csv(here::here("data/gas-price.csv"))
state_pop <- readr::read_csv(here::here("data/state-pop.csv"))
```

# Parks

Initial cleaning

```{r}
visdat::vis_dat(park_visits)
naniar::miss_var_summary(park_visits)
summary(park_visits)
# are columns completely unique for their length?
vec_size(park_visits)

completely_unique <- function(x) vec_unique_count(x) == 1
prop_unique <- function(x) (vec_unique_count(x) / vec_size(x))

# is everything totally unique?
map_lgl(park_visits, completely_unique)
map_lgl(park_visits, completely_unique) %>% any()
map_lgl(park_visits, completely_unique) %>% which()

# how unique are they?
map_dfr(park_visits, prop_unique) %>% 
  pivot_longer(cols = everything()) %>% 
  arrange(-value)
```

Let's count the number of visitors

```{r}
count(park_visits, visitors) 
```

And how many different types of patterns do we have? We can count them twice

```{r}
count(park_visits, visitors) %>% 
  count(n)
```

Lots of single visits, apparently?

What is in number of records?

```{r}
n_distinct(park_visits$number_of_records)
```

```{r}
park_visits_tidy <- park_visits %>% 
  mutate(year = parse_number(year)) %>% 
  # doesn't contain anything useful
  select(-number_of_records) %>% 
  # I want the visitor number after year
  relocate(visitors, .after = year)
```

```{r}
park_visits_tidy
```

Let's define this as a time series `tsibble`. However, to do so, we'll need to identify the index and key

- index: "year"
- key: park? 

What defines a park?

- `parkname`: character	Full park name
- `unit_code`: character	Park code abbreviation
- `unit_name`: character	Park Unit name
- `unit_type`: character	Park unit type

Are they the same length?

```{r}
park_visits_tidy %>% 
  summarise_at(.vars = vars(parkname, unit_code, unit_name),
               .funs = n_distinct)
```

Ok let's go `unit_name` to keep things simple, especially since `parkname` has a bunch of  missings. I think technically a park could be in multiple states so we won't add hierarchy here. We'll also remove the geometry column since it creates issued with `tsibble`.

```{r}
inspectdf::inspect_mem(park_visits_tidy) %>% inspectdf::show_plot()
```


# Exploring visitor counts

So the visitor counts is the thing that I'm most interest in, I think.


```{r}
ggplot(park_visits_tidy,
       aes(x = year,
           y = visitors)) + 
  geom_point()

```

outliers?

```{r}
ggplot(park_visits_tidy,
       aes(x = visitors)) + 
  geom_boxplot()

```

Where do these occur? Let's make the data a little bit smaller

```{r}
park_visits_cut <- park_visits_tidy %>% 
  select(year, 
         visitors,
         region:unit_type)

park_visits_cut %>% 
  filter(visitors > 2.5e8)
```

Do we get large numbers of visitors when there are missing values?

```{r}
library(naniar)

miss_var_summary(park_visits_cut)
gg_miss_upset(park_visits_cut)

park_visits_cut_nab <- nabular(park_visits_cut)

park_visits_cut_nab %>% 
  group_by(year_NA) %>% 
  summarise(max_visitors = max(visitors, na.rm = TRUE))

ggplot(park_visits_cut_nab,
       aes(x = year_NA,
           y = visitors)) + 
  geom_boxplot()
  
  
```

OK so based on that I think I should remove the rows that contain missing values for year and visitor_NA.

```{r}
park_visits_cut_na <- park_visits_cut %>% 
  drop_na()
```

OK now let's add on data for the state population, and the gas price, to make things a bit more interesting.  Let's quickly 

```{r}
park_visits_ts <- park_visits_cut_na %>% 
  left_join(gas_price, by = "year") %>% 
  left_join(state_pop, by = c("year", "state")) %>% 
  as_tsibble(key = unit_name,
             index = year)

park_visits_ts
```

OK, so now we get some nice spaghetti.

```{r}
gg_park_visits_spag <- ggplot(park_visits_ts,
       aes(x = year,
           y = visitors,
           group = unit_name)) + 
  geom_line()
```

Let's use brolgar to help inspect that

```{r}
library(brolgar)
gg_park_visits_spag + facet_strata(along = pop)
gg_park_visits_spag + facet_strata(along = gas_constant)
```

Let's use `keys_near` to identify those parks near the five number summary:

```{r}
park_visits_ts %>% 
  keys_near(visitors)
```

huh, filter out the zeros?

```{r}
park_visits_ts %>% 
  filter(visitors > 0) %>% 
  keys_near(visitors) 

park_visits_ts_near <- park_visits_ts %>% 
  filter(visitors > 0) %>% 
  keys_near(visitors) %>% 
  select(-visitors) %>% 
  left_join(park_visits_ts, by = "unit_name")

park_visits_ts_near
```

plot these

```{r park-near-vis}
library(stickylabeller)
ggplot(park_visits_ts_near,
       aes(x = year,
           y = visitors,
           colour = stat)) + 
  geom_line() +
  facet_wrap(~stat + unit_name, 
             scales = "free_y",
             labeller = label_glue("Park: {unit_name} \nNearest to {stat}")) + 
  theme(legend.position = "none")
```

Now plot them by the slope of year, another way of looking at it

```{r park-near-year, fig.height = 20}

park_visits_ts_lm_near <- park_visits_ts %>% 
  filter(visitors > 0) %>% 
  key_slope(visitors ~ year) %>% 
  keys_near(key = unit_name,
            var = .slope_year,
            top_n = 6) %>% 
  left_join(park_visits_ts, by = "unit_name") 

library(stickylabeller)

ggplot(park_visits_ts_lm_near,
       aes(x = year,
           y = visitors,
           colour = stat)) + 
  geom_smooth(method = "lm", se = FALSE,
              colour = "grey50",
              alpha = 0.5,
              size = 0.5) +
  geom_line() +
  facet_wrap(~stat + unit_name, 
             scales = "free_y",
             labeller = label_glue("Park: {unit_name} \nNearest to {stat} slope"),
             nrow = 10) + 
  theme(legend.position = "none")
```

Now can we add in the other variables, population size and gas to see if there's something interesting there?

```{r park-near-pop, fig.height = 20}
park_visits_lm_gas_pop <- park_visits_ts %>% 
  drop_na(pop) %>% 
  filter(visitors > 0) %>% 
  key_slope(visitors ~ year + pop + gas_constant)

park_visits_ts_lm_pop_near <- park_visits_lm_gas_pop %>% 
  keys_near(key = unit_name,
            var = .slope_pop,
            top_n = 6) %>% 
  left_join(park_visits_ts, by = "unit_name") 

library(stickylabeller)

ggplot(park_visits_ts_lm_near,
       aes(x = year,
           y = visitors,
           colour = stat)) + 
  geom_smooth(method = "lm", se = FALSE,
              colour = "grey50",
              alpha = 0.5,
              size = 0.5) +
  geom_line() +
  facet_wrap(~stat + unit_name, 
             scales = "free_y",
             labeller = label_glue("Park: {unit_name} \nNearest to {stat} slope"),
             nrow = 10) + 
  theme(legend.position = "none")

```


Are some of them always increasing?



# Gas price

```{r}
summary(gas_price)

ggplot(gas_price,
       aes(x = year,
           y = gas_constant)) + 
  geom_line()
```

# State Pop

```{r}
state_pop
```

## Reproducibility

<details><summary>Reproducibility receipt</summary>

```{r}
## datetime
Sys.time()

## repository
if(requireNamespace('git2r', quietly = TRUE)) {
  git2r::repository()
} else {
  c(
    system2("git", args = c("log", "--name-status", "-1"), stdout = TRUE),
    system2("git", args = c("remote", "-v"), stdout = TRUE)
  )
}

## session info
sessionInfo()
```

</details>
